---
title: "Lab 9"
author: 
  -  Emily Gentles
  -  Weiyi Liu
  -  Jack McCarthy
  -  Qinzhe Wang
date: "10/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(janitor)
library(reshape2)
library(kableExtra)
library(caret)
library(lattice)

knitr::opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE
)
```

```{r}
data <- read.table('BadgersFarmSurveysNoNA.txt', header=TRUE) %>%
  janitor::clean_names() %>%
  rename(no_cattle_in_buildings_yard=no_cattle_in_buidlings_yard)
```

# EDA

First a look at the distribution of the response. We can look at the mean 
__signs_in_yard__ value for each farm to get a "rate" of signs being observed.

```{r, fig.height=4}
data %>%
  group_by(farm_code_numeric) %>%
  summarise(
    n=n(),
    signs_rate=mean(signs_in_yard)
  ) %>%
  ggplot(aes(x=reorder(farm_code_numeric, -signs_rate), y=signs_rate)) +
    geom_bar(stat='identity', color='black', fill='lightcoral') +
    geom_text(aes(label=n), vjust=-0.25) +
    labs(x='Farm Code', y='Observed Signs Rate (w/ Sample Size)') +
    scale_x_discrete(guide=guide_axis(angle=45)) +
    theme_bw()
```

It seems like the signs rates vary widely across farms, so we should expect to 
see heterogeneity in our final model. Now we can look at the signs rates for 
some of the candidate predictors. First, a look at the rates for each value of 
the number of active badger settlements.

```{r, fig.height=3.5}
data %>%
  group_by(no_active_setts_in_fields) %>%
  summarise(
    n=n(),
    signs_rate=mean(signs_in_yard)
  ) %>%
  ggplot(aes(x=no_active_setts_in_fields, y=signs_rate)) +
    geom_bar(stat='identity', color='black', fill='lightblue') +
    geom_text(aes(label=n), vjust=-0.25) +
    scale_x_continuous(
      labels=as.character(unique(data$no_active_setts_in_fields)), 
      breaks=unique(data$no_active_setts_in_fields)
    ) +
    labs(x='# Active Settlements', y='Observed Signs Rate (w/ Sample Size)') +
    theme_bw()
```

There seems to be a positive relationship between the number of active 
settlements and the rate of badger signs in the yard. We may also look at the 
relationship between the number of cattle and the signs rate. To do so we will 
bin the number of cattle into 10% quantile increments and chart the rate for 
each bin.

```{r, fig.height=4}
data %>%
  mutate(cattle_bin=cut(
      no_cattle_in_buildings_yard,
      breaks=quantile(no_cattle_in_buildings_yard, (1:10)/10),
      include.lowest=TRUE
    )
  ) %>%
  group_by(cattle_bin) %>%
  summarise(
    n=n(),
    signs_rate=mean(signs_in_yard)
  ) %>%
  ggplot(aes(x=cattle_bin, y=signs_rate)) +
    geom_bar(stat='identity', color='black', fill='lightgreen') +
    geom_text(aes(label=n), vjust=-0.25) +
    scale_x_discrete(guide=guide_axis(angle=45)) +
    labs(x='# Cattle', y='Observed Signs Rate (w/ Sample Size)') +
    theme_bw()
```

Again, there seems to be a slight positive relationship between the number of 
cattle and the rate of badger sign observation. The plot below is an alternative 
view of this relationship.

```{r, fig.height=3}
data %>%
  mutate(cattle_bin=cut_interval(no_cattle_in_buildings_yard, n=25)) %>%
  group_by(cattle_bin) %>%
  summarise(
    n=n(),
    signs_rate=mean(signs_in_yard)
  ) %>%
  ggplot(aes(x=1:length(unique(cattle_bin)), y=signs_rate)) +
    geom_point() +
    geom_smooth(method='lm') +
    scale_x_discrete(guide=guide_axis(angle=45)) +
    labs(x='# Cattle (Bin Index)', y='Observed Signs Rate (w/ Sample Size)') +
    theme_bw()
```

We also have several indicator variables to consider. To investigate these, we 
can look at the accuracy in predicting the response using each indicator as a 
sole predictor.

```{r}
data %>%
  select(accessible_cattle_house_present:molasses) %>%
  sapply(., function(x) mean(x==data$signs_in_yard)) %>%
  enframe(name='Indicator', value='Accuracy') %>%
  kable(digits=4) %>%
  kable_classic(full_width=FALSE, latex_options='HOLD_position') %>%
  row_spec(c(2,8:10), bold=TRUE)
```

```{r}
confusionMatrix(
  factor(data$proteinblocks), 
  factor(data$signs_in_yard)
)
```

Note: accuracy is the proportion of the data that are predicted correctly.
The accuracy is good for some indicators because most observations are 0 for the 
response. The confusion matrix shows that the indicators do not have as much 
predictive power as it may seem from the accuracy alone.

Finally, we may look at the signs rate for each year and season combination.

```{r, fig.height=3}
data %>%
  mutate(year_season=paste0(year, '.', season)) %>%
  group_by(year_season) %>%
  summarise(
    n=n(),
    signs_rate=mean(signs_in_yard)
  ) %>%
  ggplot(aes(x=year_season, y=signs_rate)) +
    geom_bar(stat='identity', color='black', fill='plum2') +
    geom_text(aes(label=n), vjust=-0.25) +
    labs(x='Year.Season', y='Observed Signs Rate (w/ Sample Size)') +
    theme_bw()
```

2003 and 2005 seem to have similar rates, but 2004 may have an appreciably 
lower rate of badger sign observations.

```{r}
# make relevant variables into factors
model_data <- data
model_data[, c(1, 2, 10:21)] <- lapply(data[, c(1, 2, 10:21)] , factor)

model <- glmer(
  signs_in_yard ~ no_active_setts_in_fields + no_cattle_in_buildings_yard + year + (1 | farm_code_numeric),
  family='binomial',
  control=glmerControl(optimizer = "bobyqa"),
  data=model_data
)

summary(model)
dotplot(ranef(model))
```